---
title:  Visualizing Data Using ggplot2
author: |
  | Last name: Akad
  | First name: Dogan
  | Student ID: 1001386083
  | Course section: STA303H1S-L0101
date: "Jan 21st, 2017"
output: pdf_document
fontsize: 10pt
---

# \textcolor{blue}{ Q1 - Factor variables analysis}

\textbf{Question: (1 point)} How many factor variables in this data set? Use R command \textit{str(diamonds)} to find it. For each factor variable, find the one-way frequency table for it. An example of cut variable is given in the solution template.

\textbf{Answer}: We have 3 factor variables. They are cut (with 5 levels),color(with 7 levels), clarity(with 8 levels).

```{r,echo=TRUE,eval=TRUE,warning=FALSE}
library(ggplot2)
# read in the file
data(diamonds)

# check the type of variables in this data
str(diamonds)

# observations for each level of "cut" variable
table(diamonds$cut) # or summary(diamonds$cut)

# find the level frequency of "cut" varible
prop.table( table(diamonds$cut) ) # or summary(diamonds$cut)/nrow(diamonds)

# find the level frequency of "color" variable 
prop.table( table(diamonds$color) )

# find the level frequency of "clarity" variable
prop.table( table(diamonds$clarity) )

```
  
\newpage 

# \textcolor{blue}{ Q2 -Produce plot and give comment}

## Scatter plot 

```{r,echo=TRUE,eval=TRUE, warning=FALSE}
library(ggplot2)
library(plyr)
# read in the data file
data(diamonds)
qplot(carat, price, data = diamonds, colour = clarity)

```

## Comments: 
Price always increases with all the levels of clarity. Price grows close to linear when the carat has the worst clarity. As the clarity increases the price starts to grow quadratically. When the carat has the best clarity(IF) price grows the fastest. So from this plot we can see that carat clarity has a big effect on the price.


